---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
  
**CLEAR ALL VARIABLES IN WORKSPACE**  
```{r}
rm(list=ls())
```  
**LOAD PACKAGES**  
```{r}
library(lubridate)
library(tseries)
library(Hmisc)
```  
**READ DATA**  
```{r}
OSGIX <- read.csv2("C:/Users/mihal/OneDrive/OSGIX.txt",sep="",skip=20)
head(OSGIX)
tail(OSGIX)
#CHECK THAT I HAVE THE CORRECT DATA TO CONDUCT THE ANALYSIS  
independentvariables <- read.csv2("C:/Users/mihal/OneDrive/indep.txt",sep="",header=T,skip=20)
newindependentvariables <- independentvariables[1:340,]
#SELECT 340 OBSERVATIONS , EQUAL TO THE NUMBER OF THE OBSERVATIONS OF THE DEPENDANT VARIABLE OSGIX
```  
**SET UP THE TIME SERIES**  
```{r}
OSGIXts <- ts(OSGIX,frequency = 12,start = decimal_date(as.Date("1989-04-01"))) 
plot(OSGIXts, type="l", col='blue', lwd=1, main="Time Series plot of OSGIX", ylab="Mutual Fund Returns") 
#PARATIROUME VOLATILITY CLUSTERING ARA ANAMENOYME NA EXOYME ETEROSKEDASTIKOTITA LOGW TIS ENTONIS METAVLITOTITAS ARA THA XREIASTOYME GARCH MONTELA, I SEIRA KINEITAI GYRW APO TO MIDEN XWRIS NA YPARXEI TASI,DEN FAINETAI NA EXEI DOMIKES ALLAGES
```  
**SEASONALITY CHECK**  
```{r}
library(fpp2)
ggseasonplot(OSGIXts,main="Seasonal Plot of OSGIX Fund", ylab="Mutual Fund Returns")
#INDICATES THERE'S MOST LIKELY NO SEASONALITY
ggsubseriesplot(OSGIXts,main="Seasonal Plot of OSGIX Fund", ylab="Mutual Fund Returns")
#NOT TOO DIFFERENT MONTHLY AVERAGE MEANS SO THERE'S NO SEASONALITY
```  
**TRANSFORM DATA**  
```{r}
y <- OSGIX[,1]
independentvariables2 <- newindependentvariables/100
dim(independentvariables2)
X1 <-  independentvariables2[,1]
X2 <- independentvariables2[,2]
X3 <- independentvariables2[,3]
X4 <- independentvariables2[,4]
X5 <- independentvariables2[,5]
X6 <- independentvariables2[,6]
newindependentvariables2 <- cbind(X1,X2,X3,X4,X5,X6)
dim(newindependentvariables2)
xfactors <- ts(newindependentvariables2,frequency=12,start=decimal_date(as.Date("1989-04-01")))
plot(xfactors,col="blue")
```  
**CREATE AUTOCORRELATION AND PARTIAL AUTOCORRELATION PLOTS FOR OSGIX**  
```{r}
acf(y,12,main="ACF plot of OSGIX") 
#ENDEIKSEIS GIA TIN STASIMOTITA KATHWS KAI GIA TO OTI DEN YPARXEI AYTOSYSXETISI
pacf(y,12,main="PACF plot of OSGIX")
#AN YPARXEI PROBLIMA AYTOSYSXETISIS THA VALOYME TIME SERIES MONTELA
```  
**TEST FOR ALL AUTOCORRELATIONS**  
```{r}
test1<- Box.test(y,12,type="Box-Pierce")  
test1
#FAIL TO REJECT Ho,MENOYME STIN Ho DILADI DEN EXOYME AYTOSYSXETISI MEXRI TO LAG 12
test2 <- Box.test(y,12,type="Ljung-Box")
test2
#IDIA SYMPERASMATA ME DIAFORETIKI METHODO 
#WHITE NOISE OPOIO LAG ITAN EKTOS TWN DIAKEKKOMENWN GRAMMWN
```  
**CHANGE POINT CHECK**  
```{r}
library(magrittr)
library(changepoint)
y %>%
  cpt.mean(
    test.stat="Normal"
  ) %T>%
  plot(
    main="Change points for OSGIX period 1989-2017",
    xlab="Year",
    ylab="Mutual Fund Returns"
  ) %>%
  cpts.ts()
#NO CHANGE POINT DETECTED , ANAMENOMENO, I SEIRA EINAI STASIMI KAI KINEITAI GYRW APO TO 0
```  
**EXPLORE LINEAR RELATIONSHIPS**  
```{r}
rcorr(as.matrix(cbind(y,X1,X2,X3,X4,X5,X6)))
#CORRELATION COEFFΙCIENTS AND P_VALUES
```  
**SCATTERPLOT OF ALL VARIABLES**  
```{r}
pairs(cbind(y,X1,X2,X3,X4,X5,X6))
```   
  
**ESTIMATE MULTIPLE REGRESSION**  
```{r}
fitall <- lm(y~X1+X2+X3+X4+X5+X6)
summary(fitall)
# MULTIPLE R-SQUARED 89% AND R ADJUSTED SQUARED ALMOST 89% , VERY GOOD FIT!
coef(fitall)
alpha <-coef(fitall)[1]
alpha
```  
**BACKWARD ELIMINATION METHOD**  
```{r}
stepBE <- step(fitall,scope=list(lower= ~1,upper= ~X1+X2+X3+X4+X5+X6,direction="backward"))
stepBE
#KATALIGEI STO MONTELO XWRIS TIN EPEKSIGIMATIKI METAVLITI X4 
```  
**FORWARD SELECTION METHOD**  
```{r}
fitnull <- lm(y~1)
stepFS <- step(fitnull,scope=list(lower= ~1,upper= ~X1+X2+X3+X4+X5+X6),direction="forward")
#KATELIKSE STO IDIO MONTELO ME TIN BACKWARD ELIMINATION METHOD
```  
**STEPWISE SELECTION METHOD**  
```{r}
stepSS <- step(fitnull,scope=list(lower= ~1,upper= ~X1+X2+X3+X4+X5+X6),direction="both")
stepSS
#KATELIKSE STO IDIO MONTELO ME TIS 2 PROIGOYMENES METHODOYS
```  
**DIAGNOSTIC CHECKING**    
```{r}
acf(residuals(fitall),50) 
#ENDEIKSEIS OTI DEN YPARXEI AYTOSYSXETISI
```  
```{r}
pacf(residuals(fitall),50) 
#2 STA 50 EKTOS ORIWN
```  
```{r}
res1 <- Box.test(residuals(fitall),50,type="Box-Pierce")
#TEST FOR ALL AUTOCORRELATIONS 
res1
#FAIL TO REJECT Ho,MENOYME STIN Ho DILADI DEN EXOYME AYTOSYSXETISI MEXRI TO LAG 50
```  
```{r}
res2 <- Box.test(residuals(fitall),50,type="Ljung-Box")
#IDIA SYMPERASMATA ME DIAFORETIKI METHODO 
res2
#WHITE NOISE TO LAG POY ITAN EKTOS TWN DIAKEKKOMENWN GRAMMWN
```  
```{r}
acf(residuals(fitall)^2,12)
#YPARXEI PROVLIMA ETEROSKEDASTIKOTITAS STA LAGS 1,2,3,4,5,6,8,9,12
```  
```{r}
pacf(residuals(fitall)^2,12)
#YPARXEI PROVLIMA ETEROSKEDASTIKOTITAS STA LAGS 1,3,4,11
```  
```{r}
hist(fitall$residuals,col="red",xlab="Error",main="Histogram of Residuals",freq=FALSE)
lines(density(fitall$residuals))
#FAINETAI NA YPARXEI PROVLIMA KANONIKOTITAS TWN KATALOIPWN
```  
```{r}
jarque.bera.test(fitall$residuals)
#REJECT Ho AT 5% LEVEL OF SIGNIFICANCE ARA OXI KANONIKA KATANEMIMENA KATALOIPA
```  
```{r}
shapiro.test(fitall$residuals)
#DEYTEROS TROPOS GIA ELEGXO TIS KANONIKOTITAS,IDIA SYMPERASMATA
```  
```{r}
qqnorm(fitall$residuals) 
qqline(fitall$residuals)
#3oS TROPOS NA DOYME TIN MI KANONIKOTITA KATHWS KAI POY AKRIVWS OFEILEITAI TO PROVLIMA, STIS OYRES
```  
**ESTIMATING AN ARCH(1) MODEL FOR THE OSGIX TIME SERIES**  
```{r}
library(rugarch)
library(fGarch)
m1arch <- garchFit(~garch(1,0),data=y,trace=F)
summary(m1arch)
#DEN YPARXEI  PROVLIMA AYTOSYSXETISIS TWN KATALOIPWN
#PROVLIMA ETEROSKEDASTIKOTITAS
#PROVLIMA KANONIKOTITAS
#STIN PRAGMATIKOTITA DEN XREIAZETAI AYTO TO VIMA GIATI IKSERA OTI XREIAZOMAI GARCH MONTELO
```  
**ESTIMATING A GARCH(1,1) MODEL FOR THE OSGIX TIME SERIES**  
```{r}
m2garch <- garchFit(~garch(1,1),data=y,trace=F)
summary(m2garch)
#DEN YPARXEI PROVLIMA AYTOSYSXETISIS TWN KATALOIPWN
#DIORTHWTHIKE TO PROVLIMA TIS ETEROSKEDASTIKOTITAS ME TO GARCH (1,1) MONTELO
#YPARXEI PROVLIMA STIN KANONIKOTITA TWN KATALOIPWN
```  
**ESTIMATING GARCH(P,Q) MODELS FOR THE OSGIX TIME SERIES**  
```{r echo=T, results='hide',warning=FALSE}
m3garch <- garchFit(~garch(3,1),data=y,trace=F)
summary(m3garch)

m4garch <- garchFit(~garch(1,2),data=y,trace=F)
summary(m4garch)

m5garch <- garchFit(~garch(3,2),data=y,trace=F)
summary(m5garch)

m6garch <- garchFit(~garch(1,3),data=y,trace=F)
summary(m6garch)

m7garch <- garchFit(~garch(1,4),data=y,trace=F)
summary(m7garch)

m8garch <- garchFit(~garch(1,11),data=y,trace=F)
summary(m8garch)

m9garch <- garchFit(~garch(3,2),data=y,trace=F)
summary(m9garch)
#BEST MODEL BASED ON AIC IS THE GARCH(1,2)
#BEST MODEL BASED ON BIC IS THE GARCH(1,1)
```  
**RUN MULTIPLE REGRESSION MODEL + sGARCH MODEL**  
```{r}
X <- matrix(cbind(X1,X2,X3,X4,X5,X6),ncol=6)
spec <- ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,0),include.mean=TRUE,external.regressors=X),distribution.model="norm")
spec
modelres1 <- ugarchfit(spec = spec , data=y)
modelres1
#PARATIRW OTI H EPEKSIGIMATIKI METAVLITI X4 DEN EINAI STATISTIKA SIMANTIKI OPOTE TIN AFAIRW
```  
**BACKWARD ELIMINATION METHOD MANUALLY**  
```{r}  
XX <- matrix(cbind(X1,X2,X3,X5,X6),ncol=5)
spec2 <- ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,0),include.mean=TRUE,external.regressors=X),distribution.model="norm")
modelres2 <- ugarchfit(spec=spec2,data=y)
modelres2
#PARATIRW OTI H EPEKSIGIMATIKI METAVLITI X3 DEN EINAI STATISTIKA SIMANTIKI OPOTE TIN AFAIRW  
XXX <- matrix(cbind(X1,X2,X5,X6),ncol=4)
spec3 <-  ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),
mean.model=list(armaOrder=c(0,0),include.mean=TRUE,external.regressors=XXX),
distribution.model="norm")
spec3 
modelres3 <- ugarchfit(spec=spec3,data=y)
modelres3
#KATELIKSA STO MONTELO OPOY OI X1,X2,X5,X6 EPEKSIGIMATIKES METAVLITES EINAI STATISTIKA SIMANTIKES
spec4 <- ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),
mean.model=list(armaOrder=c(0,0),include.mean=TRUE,external.regressors=XXX),
distribution.model="std")
spec4
modelres4 <- ugarchfit(spec=spec4,data=y)
modelres4
#EPILEGW TIN STUDENT-T KATANOMI GIA NA SULLAVEI KALYTERA TIS PAXIES OYRES OPOTE AYTO EINAI TO OPTIMAL MONTELO,AYTO FAINETAI APO TO PLOT TWN STANDARDIZED RESIDUALS!
```  
**EXTRACT ESTIMATED COEFFICIENTS**  
```{r}
modelres4_var <- modelres4@fit$var
#SAVE THE ESTIMATED CONDITIONAL VARIANCES
modelres4_squaredres <- (modelres4@fit$residuals)^2
#SAVE THE ESTIMATED SQUARED RESIDUALS
```  
**PLOT THE SQUARED RESIDUALS AND THE ESTIMATED CONDITIONAL VARIANCES**  
```{r}
plot(modelres4_squaredres,type="l")
lines(modelres4_var,col="green",lwd=3)
```  
**RUN MULTIPLE REGRESSION + EGARCH MODEL**  
```{r}
#EGARCH MODEL IN ORDER TO CATCH THE LEVERAGE EFFECT WHICH IS NOT CAUGHT BY THE GARCH MODELS
#THE EXPONENTIAL NATURE OF THE CONDITIONAL VARIANCE IN THE EGARCH MODE CAPTURES THE EFFECT OF EXTERNAL UNEXPECTED SHOCKS ON THE PREDICTED VOLATLITY
X <- matrix(cbind(X1,X2,X3,X4,X5,X6),ncol=6)
spec5 <- ugarchspec(variance.model=list(model="eGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,0),
include.mean=TRUE,external.regressors=X),distribution.model="norm")
spec5
modelres5 <- ugarchfit(spec = spec5 , data=y)
modelres5
#SE AYTO TO MONTELO OLES OI EPEKSIGIMATIKES METAVILTES Xi EINAI STATISTIKA SIMANTIKES,AS FTIAKSOYME KAI TO PROVLIMA STIS OURES
spec6 <- ugarchspec(variance.model=list(model="eGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,0), 
include.mean=TRUE,external.regressors=X),distribution.model="std")
spec2
modelres6 <- ugarchfit(spec = spec6, data=y )
modelres6
#INTERPRETATION OF GAMMA1 : SPLITTING THE POSITIVE AND THE NEGATIVE IN ORDER TO CATCH THE LEVERAGE EFFECT
```  
**VOLATILITY OF STANDARD GARCH VS EGARCH**  
```{r}
voly_Egarch <- ts(modelres6@fit$sigma^2)
plot(voly_Egarch,lwd=2,col="blue",main="Volatility of OSGIX-ΕGARCH(1,1)")
#VOLATILITY IS THE MEASUREMENT OF RISK.HIGHER VOLATILITY LEADS TO LARGE VARIATIONS OF RETURN,HENCE HIGHER RISK
voly_Sgarch <- ts(modelres4@fit$sigma^2) 
plot(voly_Sgarch,lwd=2,col="blue",main="Volatility of OSGIX-SGARCH(1,1)")
```  
**HOW SIMILAR ARE THE SERIES**  
```{r}
cor(voly_Sgarch,voly_Egarch)
#CORRELATION IS A SINGLE NUMBER THAT DESCRIBES THE DEGREE OF RELATION BETWEEN THO VARIABLES.
```  
**PLOT THEM TOGETHER**  
```{r}
ts.plot(voly_Sgarch,voly_Egarch,col=c("green","red"),xlab="",lwd = c(1,1),
main="Time Series Plot of Volatily of SGarch vs EGarch")
legend("topright",legend=c("Standard","Exponential"),col=c("green","red"),lty=c(1,1),lwd =c(3,3))
#STANDARD GARCH HAS HIGHER HIGHS IN CERTAIN SPOTS , THERE'S A SLIGHT DIFFERENCE IN THE RESIDUAL PART OF THE TIME SERIES
```  
**DATA FOR THE FORECAST**  
```{r}
data_for_forecast <- independentvariables[341,]
#SELECT THE DATA FROM 8/2017
pred <- ugarchforecast(modelres4,n.ahead = 24,external.forecasts=data_for_forecast)
pred
#PREDICTION ABOUT THE SERIES AND THE SIGMA 24 PERIODS AHEAD
ug_f <- pred@forecast$seriesFor
plot(ug_f,type="l")
#PLOT OF THE VOLATILITY WHICH IS THE SQUARE ROOT OF THE CONDITIONAL VARIANCE
#WE OBSERVE CONSTANT VOLATILITY APPROXIMATELY 0.0010
```  
**MEAN SQUARED FORECAST ERROR**
```{r}
realdatafcOSGIX <- read.csv2("C:/Users/mihal/OneDrive/realdatafcOSGIX.txt",sep="",
col.names = "OSGIX")
realdatafcOSGIX <- realdatafcOSGIX[,1]
#OSGIX DATA FROM 8/2017 - 7/2019
i <- 1:24
Y <- realdatafcOSGIX
Y_hat <- pred@forecast$seriesFor
MSFE <- (sum(Y[i]-Y_hat[i])^2)/24
MSFE
```  
**HIT RATIO**  
```{r}
Y_pos <- Y[Y>0]
Y_pos
#THE POSITIVE VALUES OF Y
Y_neg <- Y[Y<0]
Y_neg
#THE NEGATIVE VALUES OF Y
Hit_ratio <- 19/24
Hit_ratio
#79% INDICATES A GOOD HIT RATIO
```  
**BOOTSTRAP FORECAST**  
```{r}
bootp <- ugarchboot(modelres4,method = c("Partial","Full")[1],n.ahead = 24,n.bootpred = 24)
bootp
#THE BOOTSTRAP METHOD IN THE rugarch PACKAGE IS BASED ON RESAMPLING STANDARDIZED RESIDUALS FROM THE EMPIRICAL DISTRIBUTION OF THE FITTED MODEL TO GENERATE FUTURE REALIZATIONS OF THE SERIES AND SIGMA (THE SQUARE ROOT OF THE VARIANCE)
#USING plot(bootp) WE CAN OBTAIN THE ESTIMATED SERIES AS WELL AS SIGMA STANDARD ERROR PLOTS
```  
**FORECAST USING MULTIPLE LAYER PERCEPTRON**  
```{r}
library(nnfor)
set.seed(1)
#SET RANDOM SEED FOR REPRODUCIBILITY IN PROCESSES LIKE SPLITTING THE DATA,ANY NUMBER WILL DO
ynew <- read.csv2("C:/Users/mihal/OneDrive/OSGIX.txt",header=T,sep="",skip=29)
ynew2 <- ynew[1:324,]
df <- ts(ynew2,frequency = 12,start = decimal_date(as.Date("1990-01-01")))
#SPLITTING DATA INTO A TRAINING DATA SET AND A TESTING DATA SET
trn <- df %>% window(end = c(2014,12))
trn
trn %>% 
  plot(
    main = "OSGIX SERIES PLOT",
    ylim = c(-0.3,0.2)
  )
tst <- df %>% window(start=2015)
tst
fit1 <- trn %>% mlp()
#DEFAULT MLP MODEL HAS 5 HIDDEN NODES
pred1 <- fit1 %>% forecast(h=48)
#BLUE => FORECAST 
pred1
pred1 %>% plot()
tst %>% lines(lwd=2,col="red")
#RED => REAL DATA
```  
  
**SECOND METHOD TO FORECAST USING MLP**  
```{r,eval=FALSE}
#FIT MLP MODEL WITH NUMBER OF HIDDEN NODES DETERMINED BY 20% VALIDATION
fit2 <- trn %>% mlp(hd.auto.type = "valid")
pred2 <- fit2 %>% forecast(h=48)
pred2
pred2 %>% plot()
tst %>% lines(lwd=2,col="red")
# 13 MINUTES PROCEDURE
```  
**THIRD METHOD TO FORECAST USING MLP**  
```{r,eval=FALSE}
#FIT MLP MODEL WITH NUMBER OF HIDDEN NODES DETERMINED BY 5 FOLD CROSS VALIDATION
fit3 <- trn %>% mlp(hd.auto.type = "cv")
pred3 <- fit3 %>% forecast(h=48)
pred3
pred3 %>% plot()
tst %>% lines(lwd=2,col="red")
#LONGER PROCEDURE
```  
**FORECASTING USING GENERALIZED REGRESSION NEURAL NETWORKS (GRNN)**  
```{r}
library(tsfgrnn)
pred <- grnn_forecasting(y,h = 24)
#THE LAGS USED AS AUTOREGRESSIVE VARIABLES ARE SET WITH THE LAGS PARAMETER.IF NULL(THE DEFAULT) THE LAGS ARE SELECTED IN A FAST,HEURISTIC WAY.
pred$prediction
plot(pred,main="FORECASTING OSGIX USING GRNN")
library(ggplot2)
autoplot(pred)
```  
**NEURAL NETWORK AUTOREGRESSION (OR NNAR MODEL)**  
```{r}
#FEED-FORWARD NEURAL NETWORKS WITH A SINGLE HIDDEN LAYER AND LAGGED INPUTS FOR FORECASTING UNIVARIATE TIME SERIES
library(lattice)
library(caret)
y <- OSGIX[,1]
fit2 <- nnetar(y,repeats = 100,size = 4)
plot(forecast(fit2,h=24,main="FORECASTING OSGIX USING NNAR MODEL"))
points(1:length(y),fitted(fit2),type="l",col="red")
print(forecast(fit2))
#POINT FORECAST
```  




